---
title: "Preprocessing of scRNASeq datasets"
author: "Monika Krzak"
date: "November 5, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r cars, echo=FALSE, include=FALSE}
library(scater)
```

## Dataset

```{r, echo=FALSE}
sub('\\.Rdata$', '',  unlist(strsplit(data[i], "/"))[length(unlist(strsplit(data[i], "/")))] )
```

## **** Quality control **** 

### Filter ERCC spike-ins 
TRUE - spike-in\
FALSE - not spike-in

```{r, echo=FALSE}
table(grepl("^ERCC-", rownames(sce)))
if(length(unique((grepl("^ERCC-", rownames(sce)))))==2){
    sce=sce[-which(grepl("^ERCC-", rownames(sce))),]
}

```

### Filter duplicated genes
TRUE - duplicated\
FALSE - not duplicated


```{r, echo=FALSE}
table(duplicated(rowData(sce)$feature_symbol))
sce <- sce[!duplicated(rowData(sce)$feature_symbol), ]
```

### Filter not expressed genes
FALSE - not expressed genes\
TRUE - expressed genes

```{r, echo=FALSE}
keep_genes <- rowSums(counts(sce) > 0) > 0
sce <- sce[keep_genes, ]
table(keep_genes)
```

### Total counts

```{r, echo=FALSE, include=FALSE, warning=FALSE}
sce <- calculateQCMetrics(sce)
```

```{r, echo=FALSE, warning=FALSE, comment=FALSE}
hist(sce$total_counts,breaks = 100)
```

### Number of detected genes

```{r, echo=FALSE}
hist(sce$total_features_by_counts,breaks = 100)
```

### Filter outlier cells
Filter cells with total counts and number of detected features more than 3 median absolute deviations (MADs) below the median across all the cells (on log scale). 

TRUE - outlier\
FALSE - not outlier

```{r, echo=FALSE, warning=FALSE}
colData(sce)$libsize.drop <- isOutlier(sce$total_counts, nmads = 3, type = "lower", log = TRUE)
colData(sce)$feature.drop <- isOutlier(sce$total_features_by_counts, nmads = 3, type = "lower", log =TRUE)
sce <- sce[, !(sce$libsize.drop | sce$feature.drop)]
table(sce$libsize.drop | sce$feature.drop)
```

### Dimension after quality control

```{r, echo=FALSE}
paste0("Number of genes: ", dim(sce)[1])
paste0("Number of cells: ", dim(sce)[2])
```

### Save qc dataset 

```{r, echo=FALSE}
save(sce, file=paste0("Rdata_qc/", unlist(strsplit(data[i], "/"))[length(unlist(strsplit(data[i], "/")))]))
```

## **** First preprocessing: filtering after quality control ****

### Filter genes 
Filter genes with average expression count (adjusted by library size) equal to 0

```{r, echo=FALSE}
rowData(sce)$ave.counts <- calcAverage(sce, exprs_values = "counts", use_size_factors=FALSE)
to.keep <- rowData(sce)$ave.counts > 0
sce <- sce[to.keep,]
```

### Dimension after filtering

```{r, echo=FALSE}
paste0("Number of genes: ", dim(sce)[1])
paste0("Number of cells: ", dim(sce)[2])
paste0("Number of groups: ",length(unique(sce$Group)))


```

### Save filtered and qc dataset 

```{r, echo=FALSE}
save(sce, file=paste0("Rdata_qc_filt/", unlist(strsplit(data[i], "/"))[length(unlist(strsplit(data[i], "/")))]))
```

## **** Second preprocessing: normalization after filtering and quality control ****

### Normalize dataset

```{r, echo=FALSE}
library(scran)
set.seed(100)
if(dim(sce)[2]<500){
 min.size=50   
} else {
 min.size=200   
}

clusters <- quickCluster(sce, min.size=min.size, min.mean=0.1, method="igraph")
sce <- computeSumFactors(sce, cluster=clusters, min.mean=0.1)
summary(sizeFactors(sce))
sce <- scater::normalize(sce, return_log = FALSE)
counts(sce) <- normcounts(sce) #assign normalized counts to slot counts

```

### Save normalized, filtered and qc dataset 

```{r, echo=FALSE}
save(sce, file=paste0("Rdata_qc_filt_norm/", unlist(strsplit(data[i], "/"))[length(unlist(strsplit(data[i], "/")))]))
```

### PCA 
PCA on log2 transformed normalized dataset
```{r, echo=FALSE}
logcounts(sce) <- log2(normcounts(sce)+1)
scater::plotPCA(sce, colour_by="Group")
```





















